Calculating Entropies With Nested Sampling

### 2017-02-13 paper/ms.tex
Some final numbers
New results

Letting $\tau = \log_{10} T$, and treating
$A$ and $\phi$ are nuisance parameters, I calculated the entropies
of $\tau$, $\d$, and $(\tau, \d)$.
These results were computed using a tolerance of 0.001 for every quantity.
The results were
\begin{align}
H_{\tau}    &= 7.337 \pm 0.028 \\
H_{\d}      &= 905.9 \pm 1.9   \\ % Interim result. Check gave 903.3 +- 10.9
H_{\tau,\d} &= 902.2 \pm 1.7      % Interim result only
\end{align}

@@

To convert these to differential entropies, we need to add the appropriate
log-volume to each. These are $-6.215$, $-790.32$, and
$-796.53$ respectively, so the differential entropies are:
\begin{align}
H'_{\tau}     &= 1.122 \pm 0.028 \\
H'_{\d}       &= 115.5 \pm 1.9   \\
H'_{\tau, \d} &= 105.7 \pm 1.7
\end{align}
The mutual information is equal to $H_{\tau} + H_{\d} - H_{\tau, \d}$. The
uncertainties in each of these three quantities is quite large, and would lead
to a large uncertainty on $I_{\tau, \d}$ if computed naively. To reduce this
uncertainty, we computed the three entropies using a common random number
generator for the reference pairs. This is an example of the
{\em common random numbers} technique of variance reduction.
The resulting mutual information is
$I_{\tau, \d} = 11.1 \pm 1.1$ nats {\todo (interim result only)}.

The computational resources needed to compute these quantities are quite large,
though the algorithm can be parallelised trivially. However, the problem
being solved is not trivial. The mutual information of $\x$ and
$\d$ is
\begin{align}
I_{\x, \d} &= \int p(\x, \d) \log\frac{p(\x, \d)}{p(\x)p(\d)} \, d\x \, d\d.
\end{align}
However, $\x$ is one of several parameters. The others, $A$, and $\phi$,
were implicitly integrated out. Denoting the nuisance parameters by
$\n$, the mutual information that has been calculated is
\begin{align}
I_{\x, \d} &= \int p(\x, \d)
  \log\frac{\int p(\x, \n, \d) \, d\n }
           {p(\x) \int p(\x', \n)p(\d | \x', \n) \, d\x' \, d\n}
  \, d\x \, d\d.
\end{align}
It should not be considered surprising that this is a challenging task.

\subsection{Uneven observations}
The results for the uneven observing schedule are
\begin{align}
I_{\x, \d} &= 12.09 \pm 0.82
\end{align}
which appears to be higher than under the even observing schedule,
though the uncertainties
are not small enough to be certain of this.

>>>>

Letting $\tau = \log_{10} T$, and treating
$A$ and $\phi$ are nuisance parameters, I calculated the entropies
of $\tau$, $\d$, and $(\tau, \d)$.
These results were computed using a tolerance of 0.001 for every quantity.
The results were
\begin{align}
H_{\tau}    &= 7.337 \pm 0.028 \\
H_{\d}      &= 903.8 \pm 1.1   \\
H_{\tau,\d} &= 900.8 \pm 1.1      % Interim result only
\end{align}

@@

To convert these to differential entropies, we need to add the appropriate
log-volume to each. These are $-6.215$, $-790.32$, and
$-796.53$ respectively, so the differential entropies are:
\begin{align}
H'_{\tau}     &= 1.122 \pm 0.028 \\
H'_{\d}       &= 115.5 \pm 1.9   \\
H'_{\tau, \d} &= 105.7 \pm 1.7
\end{align}
The mutual information is equal to $H_{\tau} + H_{\d} - H_{\tau, \d}$. The
uncertainties in each of these three quantities is quite large, and would lead
to a large uncertainty on $I_{\tau, \d}$ if computed naively. To reduce this
uncertainty, we computed the three entropies using a common random number
generator for the reference pairs. This is an example of the
{\em common random numbers} technique of variance reduction.
The resulting mutual information is
$I_{\tau, \d} = 10.35 \pm 0.61$ nats.

The computational resources needed to compute these quantities are quite large,
though the algorithm can be parallelised trivially. However, the problem
being solved is not trivial. The mutual information of $\tau$ and
$\d$ is
\begin{align}
I_{\tau, \d} &= \int p(\x, \d) \log\frac{p(\x, \d)}{p(\x)p(\d)} \, d\x \, d\d.
\end{align}
However, $\x$ is one of several parameters. The others, $A$, and $\phi$,
were implicitly integrated out. Denoting the nuisance parameters by
$\n$, the mutual information that has been calculated is
\begin{align}
I_{\x, \d} &= \int p(\x, \d)
  \log\frac{\int p(\x, \n, \d) \, d\n }
           {p(\x) \int p(\x', \n)p(\d | \x', \n) \, d\x' \, d\n}
  \, d\x \, d\d.
\end{align}
It should not be considered surprising that this is a challenging task.

\subsection{Uneven observations}
The results for the uneven observing schedule are as follows:
\begin{align}
H'_{\tau}     &= 1.122 \pm 0.028 \\
H'_{\d}       &= 106.8 \pm 1.1   \\
H'_{\tau, \d} &= 97.0 \pm 1.0
I_{\tau, \d} &= 10.9 \pm 0.4
\end{align}
which appears to be higher than under the even observing schedule,
though the uncertainties
are not small enough to be certain of this.
